---
title: "Getting to know the SingleCellExperiment Object"
author:
  - name: Michael Lynch
    affiliation: School of Medicine, University of Limerick
    email: michael.lynch@ul.ie
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
vignette: >
  %\VignetteIndexEntry{singlecellexperiment object}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

````{r message=F, warning=F}

library(SingleCellExperiment)
library(scuttle)
library(muscData)

````

## Bioconductor

[Bioconductor](https://www.bioconductor.org/) is a project which aims to develop and share open-source software for precise and repeatable analysis of biological data.
They foster an inclusive and collaborative community of developers and data scientists.

## Why Bioconductor?

- Analysis workflows are long and complex, an ecosystem for R packages for genomics data analysis facilitates reuse.
- Common data structures allow interoperability between packages.
- Package build system automates testing, ensures compatability, and improves reprehensibility.
- Community events and forums improve collaboration.

You can:

- Use high-quality, interoperable, and well-documented Bioconductor R packages.
- Develop and [contribute](https://contributions.bioconductor.org/bioconductor-package-submissions.html) your own R packages for genomics analysis.

## Accessing Bioconductor packages

You might typically install a package from CRAN through `install.packages()`. 
Bioconductor packages are hosted elsewhere and have a different mechanism for installing.

First, we'll install BiocManager from CRAN, and we can then use BiocManager to install packages from Bioconductor.

````{r eval=F}

install.packages('BiocManager')
BiocManager::install('SingleCellExperiment')
````

As well as hosting, testing, and distributing packages, Bioconductor also ships some core infrastructure which improves interoperability between packages.
These usually come in the form of data structures for storing different types of data.
In this workshop, we'll use the SingleCellExperiment class.

````{r}

sce<-Kang18_8vs8()

````

The SingleCellExperiment object keeps the different types of data relevant to our analysis neatly organised in one object.
Each data point is associated with a feature and an observation, but we might also want to keep other information, for example: 

- each feature (gene) can be named using different conventions, or have different aliases.
- each observation (cell) might need its patient ID, sequencing batch, quality metrics or cluster ID attached.

````{r}

class(sce)

sce

counts(sce)[1:4,1:4]

head(colData(sce))

head(rowData(sce))

````

For convenience, column data (data describing the cells) can be accessed using the '$' operator.

````{r}

table(sce$cluster)

table(sce$ind,sce$stim)

````


We can also store transformations of the raw data, e.g. normalised counts.

````{r}

assays(sce)
sce<-logNormCounts(sce)
assays(sce)

````
Aaaand we can also store embeddings of the data, such as from PCA, TSNE or UMAP.

````{r}

reducedDimNames(sce)

````

Finally, we can store alternative experiments, or modalities, which were measured on the same cells.
For example, [CITE-seq](https://emea.illumina.com/techniques/sequencing/rna-sequencing/cite-seq.html) data measures gene and surface protein expression simultaneously.

Why might we want to treat these separately rather than appending them to the main counts?

````{r}

altExpNames(sce)

````


