---
title: "scRNA-seq analysis workshop"
author: Michael Lynch^[michael.lynch@ul.ie]
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{scRNA-seq workshop}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

````{r}

library(dittoSeq)
library(scuttle)
````

# Single-cell RNA-seq in Bioconductor workshop

Authors:
    Michael Lynch^[University of Limerick]
    <br/>
Last modified: 20th August, 2026.

## Overview

### Description

This workshop introduces the basics of single-cell RNA sequencing (gene expression) and a minimal analysis workflow to students with a background in data science, machine learning, and statistics.

Along with the topic of your workshop, include how students can expect
to spend their time. For the description may also include information
about what type of workshop it is (e.g. instructor-led live demo, lab,
lecture + lab, etc.). Instructors are strongly recommended to provide
completely worked examples for lab sessions, and a set of stand-alone
notes that can be read and understood outside of the workshop.

### Pre-requisites

* Basic knowledge of R syntax
* Basic machine learning and statistics
* Some previous data science/analysis experience

### Useful resources

- https://bioconductor.org/books/release/OSCA/
- https://bioconductor.org/books/release/OSTA/

### Participation

Interactive discussion, follow-along coding, reflection.

### _R_ / _Bioconductor_ packages used

scran, scater

### Time outline

2 hours workshop??

| Activity                     | Time |
|------------------------------|------|
| Analysis plan                | 30m  |
| scRNA-seq workflow           | 60m  |
| Reflection & Best Practices  | 30m  |

### Workshop goals and objectives

List "big picture" student-centered workshop goals and learning
objectives. Learning goals and objectives are related, but not the
same thing. These goals and objectives will help some people to decide
whether to attend the conference for training purposes, so please make
these as precise and accurate as possible.

*Learning goals* are high-level descriptions of what
participants will learn and be able to do after the workshop is
over. *Learning objectives*, on the other hand, describe in very
specific and measurable terms specific skills or knowledge
attained. The [Bloom's Taxonomy](#bloom) may be a useful framework
for defining and describing your goals and objectives, although there
are others.

### Learning goals

Some examples:

* Understand single-cell data importance and application.
* Apply data science principles to analyse new data types.
* Understand the role of Bioconductor infrastructure in research space.

### Learning objectives

* Discuss developing analysis plan for single-cell data.
* create xyz plots
* evaluate xyz data for artifacts

## Workshop

Divide the workshop into sections (`## A Section`). Include
fully-evaluated _R_ code chunks. Develop exercises and solutions, and
anticipate that your audience will walk through the code with you, or
work on the code independently -- do not be too ambitious in the
material that you present.


## Data Loading

````{r}

library(muscData)
sce<-Kang18_8vs8()

````

## Quality Control

````{r}

is.mito <- grep('^MT-',rownames(sce))
rownames(sce)[is.mito]
rowSums(counts(sce))[is.mito]

sum(rowSums(counts(sce))==0)

library(scuttle)
df <- perCellQCMetrics(sce, subsets=list(Mito=is.mito))
summary(df$sum)

hist(df$subsets_Mito_sum)

sce$qc_sum<-df$sum
sce$qc_detected<-df$detected


````

````{r}

dittoPlot(sce,c('qc_sum','qc_detected'),group.by='stim')

````
## Dimension Reduction



## Clustering



## Annotation


## Differential composition





````{r}

sessionInfo()

````
