---
title: "singlecellexperiment_object"
author: Michael Lynch^[michael.lynch@ul.ie]
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{singlecellexperiment object}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



````{r}
library(scuttle)
library(muscData)
sce<-Kang18_8vs8()

````

The SingleCellExperiment object keeps the different types of data relevant to our analysis neatly organised in one object.
Each data point is associated with a feature and an observation, but we might also want to keep other information, for example: 
- each feature (gene) can be named using different conventions, or have different aliases.
- each observation (cell) might need its patient ID, sequencing batch, quality metrics or cluster ID attached.

````{r}

sce

counts(sce)[1:4,1:4]

head(colData(sce))

head(rowData(sce))

````
For convenience, column data (about the cells) can be accessed using the '$' operator.

````{r}

table(sce$cluster)

table(sce$ind,sce$stim)

````


We can also store transformations of the raw data, e.g. normalised counts.

````{r}

assays(sce)
sce<-logNormCounts(sce)
assays(sce)
````
We can also store embeddings of the data, such as from PCA, TSNE or UMAP.

````{r}

reducedDimNames(sce)

````

Finally, we can store alternative experiments, or modalities, which were measured on the same cells.
For example, [CITE-seq](https://emea.illumina.com/techniques/sequencing/rna-sequencing/cite-seq.html) data measures gene and surface protein expression simultaneously.

Why might we want to treat these separately?

````{r}

altExpNames(sce)

````


